---
title: "Unveiling Housing Dynamics in King County, WA "
author: "Gizem Gulsiye Guleli & Duae Marriam"
date: "2023-12-19"
output: html_document
---

```{r Load necessary libraries }

library(readr) #read daatset
library(dplyr) 
library(ggplot2) 
library(plotly)
library(reshape2)
library(ggpubr)  # For ggarrange function
# Load required libraries
library(tidyverse)
library(ggrepel)
library(ggalt)

```


##  Summary

The goal of this project is to leverage advanced visualization techniques in R to analyze house prices in King County, Washington. The dataset, obtained from Kaggle, comprises 21 variables and 21,613 observations, spanning the period from 02 May 2014 to 27 May 2015

## Objectives

Develop advanced visualizations to explore relationships between variables and understand patterns in house sales data. Try to identify and interpret factors contributing to the value of houses.


## Data Overview

### Dataset

```{r  Read the dataset}

house_data <- read.csv("kc_house_data.csv",header = TRUE, sep = ",")

```

*Source:* Kaggle
*Link:* https://www.kaggle.com/datasets/shivachandel/kc-house-data/data
*Variables:* 21 (id, date, price, bedrooms, bathrooms, sqft_living, sqft_lot, floors, waterfront, view, condition, grade, sqft_above, sqft_basement, yr_built, yr_renovated, zipcode, lat, long, sqft_living15, sqft_lot15)
*Observations:* 21.613
*Period:* 02 May 2014 to 27 May 2015
*Geographic coverage:* King County, including Seattle

#### Structure of Dataset

```{r  Structure of dataset}

str(house_data)

```


The dataset consists of 21,613 observations (rows) and 21 variables (columns).
The variables include information such as id, date, price, bedrooms, bathrooms, sqft_living, sqft_lot, floors, waterfront, view, condition, grade, sqft_above, sqft_basement, yr_built, yr_renovated, zipcode, lat, long, sqft_living15, and sqft_lot15.

The id variable appears to be a unique identifier for each observation. All variables structured as numeric/integer except the date variable which is currently stored as a character type. It might be useful to convert it to a date type for time-related analyses. So we will first convert the date variable from character to a date type.

It's important to note that while all variables may be structured as numeric, certain variables, despite their numeric representation, hold categorical significance. These categorical variables are essentially numerically coded to represent different categories or levels within the dataset. This nuance is crucial to consider when interpreting and analyzing the data.


```{r Convert 'date' to Date type}

house_data$date <- as.Date(house_data$date, format = "%Y%m%dT%H%M%S")

# Verify the changes
str(house_data$date)

```

#### Summary Statistics 

```{r Display summary statistics }

summary(house_data)

```


**id**
   - The `id` variable represents a unique identifier for each home sold.

**date**
   - The `date` variable, contains information about the date of the house sale and spans from May 2, 2014, to May 27, 2015.

**price:**
   - The `price` variable is the **dependent variable**   and shows a wide range, with the minimum house price at $75,000 and the maximum at $7,700,000.
   - The median house price is $450,000, and the mean is $540,088.

**bedrooms and bathrooms:**
   - The variables related to the number of bedrooms and bathrooms  (0.5 accounts for a room with a toilet but no shower) exhibit varying ranges and distributions.
   - The number of bedrooms ranges from 0 to 33, with a mean of approximately 3.37.
   - The number of bathrooms ranges from 0 to 8, with a mean of approximately 2.12.

**sqft_living and sqft_lot:**
   - These variables represent the size of houses.
   - `sqft_living` reflects the Square footage of the apartments interior living area, ranging from 290 to 13,540 square feet, with a mean of 2080.
   - `sqft_lot` represents the lot size, ranging from 520 to 1,651,359 square feet, with a mean of 15,107.

**floors**
 -The `floors` variable is represents the levels of the houses. The majority of houses have 1 or 1.5 floors.
   - Notably, there seems to be a common occurrence of houses with 1.5 floors, while the mean is approximately 1.494.
   - This suggests that many houses have a split-level design or additional space on an upper level, contributing to the fractional floor values.

**waterfront:**  
   - The `waterfront` variable is a dummy variable mostly 0 , represents the property has no waterfront view and 1 for with waterfront.

**view and condition:**  
   - `view` represents the overall view rating (0 to 4)  with a mean of 0.23 
   -`condition` represents the overall condition rating (0 to 5) with a mean of 3.41 for `condition`.

**grade:**
   - `grade` represents the overall grade given to the housing unit and ranges from 1 to 13 where 1-3 falls short of building construction and design, 7 has an average level of construction and design, and 11-13 have a high quality level of construction and design.
   
**sqft_above, and sqft_basement:**
   - `sqft_above` and `sqft_basement` show the square footage above ground and is below ground level¶ (in the basement), respectively.

**yr_built, yr_renovated:**
   - Houses were built between 1900 and 2015 (`yr_built`), with the majority built in the mid to late 20th century.
   - `yr_renovated` indicates the last renovation year, with a mean of 84.4 and many zero values, suggesting no renovations.


**Geographical Information (lat, long, zipcode):**
   - `lat` and `long` provide latitude and longitude information of house locations, respectively.
   - `zipcode` represents the zip code of the house location.
   

**sqft_living15 and sqft_lot15:**
   - `sqft_living15` and `sqft_lot15` indicate the living room and lot size in 2015, reflecting potential renovations or changes. (? some sources mention it differently and main source couldnt find !!!)


These summary statistics provide an overview of the distribution and characteristics of each numeric variable in the dataset, with a specific focus on understanding the relationships with the **dependent variable, 'price.'**

**Missing Values:**

   - Most variables in the dataset have complete data; however, it's worth noting that `sqft_above` has two missing values (NA's).
   - Given the small number of missing values (only two observations) in relation to the overall dataset size, we have decided to remove these specific observations. This decision is based on considering the number of observations and the minimal impact on the overall analysis.
   - Removing these observations ensures that the dataset remains largely complete and is a reasonable approach in this context.

```{r Cleaning}

# Remove observations with missing values in 'sqft_above'
house_data2 <- house_data[complete.cases(house_data$sqft_above), ]

```

### Data Exploration

In organizing our variables by type, we enhance the precision of our analysis and visualization methods. This thoughtful categorization enables us to apply tailored techniques to each variable type, ensuring more insightful and nuanced exploration of the dataset.

```{r Define variables according to their types}

# All variables
all_vars <- house_data2[, c("price", "bedrooms", "bathrooms", "sqft_living", "sqft_lot","floors", "waterfront", "view", "condition", "grade", "sqft_above", "sqft_basement", "yr_built", "yr_renovated", "zipcode", "lat", "long", "sqft_living15", "sqft_lot15")]

# Categorical Variables
cat_vars <- c("waterfront", "view", "condition", "grade")

# Continuous Numeric Variables
cont_vars <- c("price", "sqft_living","sqft_living15","sqft_lot", "sqft_lot15","sqft_above", "sqft_basement")

# Discrete Numeric Variables
disc_vars <- c("bedrooms", "bathrooms", "floors")

# Date Variables
date_vars <- c("date", "yr_built", "yr_renovated")

# Geographical Variables
geo_vars <- c("lat", "long", "zipcode")

```




### Exploratory Data Analysis (EDA)

Utilize various R packages (e.g., ggplot2, plotly) for data exploration.
Conduct correlation analysis, distribution analysis.


```{r Set options}
#  to display numeric values without scientific notation and with more digits
options(scipen = 999, digits = 9)
```

#### Distribution of Contunious Numeric Variables

```{r  Distribution of Contunious Variables}

# Set up a layout grid
par(mfrow = c(3, 2), mar = c(4, 4, 2, 1))  # Adjust margins for better appearance

# Create histograms for  numeric variables
for (cont in cont_vars) {
  # Determine appropriate bin width based on the range and number of observations
  bin_width <- (max(house_data2[[cont]]) - min(house_data2[[cont]])) / sqrt(length(house_data2[[cont]]))
  
  # Create histogram with scaled x-axis
  hist(house_data2[[cont]], main = paste("Distribution of", cont), xlab = cont, col = "skyblue", breaks = seq(min(house_data2[[cont]]), max(house_data2[[cont]]) + bin_width, bin_width))

  # Add smoother distribution line
    density_curve <- density(house_data2[[cont]], bw = "nrd0")
    lines(density_curve$x, density_curve$y * bin_width * length(house_data2[[cont]]), col = "red", lwd = 2)
    
  # Add normal distribution line
  mu <- mean(house_data2[[cont]])
  sigma <- sd(house_data2[[cont]])
  x <- seq(min(house_data2[[cont]]), max(house_data2[[cont]]), length = 100)
  y <- dnorm(x, mean = mu, sd = sigma) * bin_width * length(house_data2[[cont]])
  lines(x, y, col = "blue", lwd = 2)  
  
  # Identify potential outliers using a boxplot
  boxplot(house_data2[[cont]], main = paste("Boxplot of", cont), col = "lightblue", border = "black", horizontal = TRUE)
  }

# Reset the plotting layout
par(mfrow = c(1, 1))
```

The visual inspection of the plots suggests that distribution of the variables are skewed right/ non-normal distributions with a considerable number of outliers. Given the context of the dataset, where very luxurious or unique properties may contribute to these extreme values, it is justifiable to observe such outliers.

Instead of removing or transforming these outliers, a more suitable strategy might involve employing robust statistical methods to handle outliers problem for further analysis. Robust methods are designed to be less sensitive to extreme values, allowing for a more reliable analysis that acknowledges the presence of these high-end properties without disproportionately impacting the results.




```{r}
# Load the required libraries
library(ggplot2)
library(gridExtra)

# Set up a layout grid
par(mfrow = c(3, 2), mar = c(4, 4, 1, 1))

# Create an empty list to store plots
plots_list <- list()

# Loop through continuous variables and create histograms
for (var in cont_vars) {
  # Calculate binwidth 
  bin_width <- (max(house_data2[[var]]) - min(house_data2[[var]])) / sqrt(length(house_data2[[var]]))
  
  # Create histogram with smooth distribution line
  plot <- ggplot(house_data2, aes(x = !!sym(var))) +
    geom_histogram(aes(y = ..density..), binwidth = bin_width, fill = "lightgreen", alpha = 0.7) +
    stat_density(geom = 'line', color = 'red', size = 1) +
    stat_function(fun = dnorm,
                  color = 'blue',
                  size = 1,
                  geom = 'line',
                  args = list(mean = mean(house_data2[[var]]), 
                              sd = sd(house_data2[[var]]))) +
    labs(title = paste("Distribution of", var), x = var, y = "Frequency") +
    theme_minimal()

  # Add the plot to the list
  plots_list[[var]] <- plot

}

# Reset the plotting layout
par(mfrow = c(1, 1))

# Arrange the plots together using grid.arrange
grid.arrange(grobs = plots_list, ncol = 2)


```

#### Distribution of Discrete Numeric Variables

```{r Distribution of Discrete Numeric Variables}

# Visualize the Distribution 
for (disc in disc_vars) {
  # Convert discrete numeric variables to factors
  house_data2[[disc]] <- as.factor(house_data2[[disc]])
  
  # Create bar plots for discrete numeric variables
  bar_plot <- ggplot(house_data2, aes(x = !!sym(disc), fill = !!sym(disc))) +
    geom_bar(position = "dodge", fill = "skyblue", alpha = 0.7, width = 0.7) +
    labs(title = paste("Distribution of", disc), x = disc, y = "Count") +
    theme_minimal() +
    geom_text(stat = "count", 
              aes(label = ifelse(round(after_stat(count)/sum(after_stat(count)), 3) > 0, scales::percent(round(after_stat(count)/sum(after_stat(count)), 3)), "")),
              position = position_dodge(0.7), vjust = -0.3, size= 3) +  # Add percentage labels only if count > 0
    theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels
  
  # Display the plot
  print(bar_plot)
}

```
Bedrooms :

The distribution of bedrooms in the dataset reveals a clear preference for houses with 3 bedrooms, constituting nearly half of the entries (45.5%). 4-bedroom homes follow closely with 31.8%, and 2 and 5-bedroom configurations are also prevalent, making up 12.8% and 7.4%. However, 0-bedroom and 1-bedroom houses have notably lower percentages with approximately 0.1% and 0.9%, respectively. The distribution is positively skewed, with a peak around 3 bedrooms.

Bathrooms Configuration:

The dataset showcases a diverse distribution of bathrooms. Houses with 2.5 bathrooms are most common, representing 24.9%. Additionally, 1 bathroom and 1.75 bathrooms are prevalent at 17.8% and 14.1%, respectively. The distribution exhibits multiple peaks, suggesting a variety of bathroom count configurations in the dataset.

Floor Counts:

When considering the number of floors, Houses wıth 1 floor are predominant, making up 49.4% of the dataset. 2 floors houses follow closely at 38.1%, with 1.5 floors representing 8.8%. The distribution is skewed towards fewer floors, with a sharp decline for houses with more than 2 floors.

Note on 0 Values:

In the context of houses requiring bedrooms and bathrooms, the presence of 0 values in these categories may indicate missing or incomplete data. It's uncommon for a house to have zero bedrooms or bathrooms. Investigating and addressing the reasons behind these zero values is crucial for ensuring the quality and accuracy of the dataset, as well as the reliability of any analyses conducted.

```{r}
# Visualize the Distribution of Categorical Variables 
for (cat in cat_vars) {
  # Create more advanced bar plots for categorical variables
  bar_plot <- ggplot(house_data2, aes(x = factor(!!sym(cat)), fill = factor(!!sym(cat)))) +
    geom_bar() +
    geom_text(stat = "count", aes(label = scales::percent(round(after_stat(count)/sum(after_stat(count)), 3))), vjust = -0.5) +  # Add percentage labels
    labs(title = paste("Distribution of", cat), x = cat, y = "Count") +
    theme_minimal() +
    theme(legend.position = "none")
  
  # Display the plot
  print(bar_plot)
}

```

Out of all observations, only 1 percent of houses are located on the waterfront.
Additionally, the majority of houses (90.2%) have a view score of 0. Among the remaining view scores, 4.5% have a score of 2, while scores of 1 and 4 each account for 1.5%. The remaining 2.4% of houses have a view score of 3.

The majority of houses in the dataset are in good to average condition. Approximately 91.7% of houses fall within Condition 3, indicating that a significant portion of the properties is well-maintained.Condition 4 homes represent 26.3%, suggesting a sizable proportion of houses are in better-than-average condition. Meanwhile, Condition 5 homes, which likely denote excellent condition, constitute 7.9% of the dataset.

The distribution of grades reflects a diverse range of housing quality. A significant portion of houses falls within Grade 7 (41.6%) and Grade 8 (28.1%), indicating properties with a higher level of construction and design. Grades 9 and 10 together contribute 17.3%, highlighting a considerable proportion of houses with superior construction and design quality.The dataset includes a limited number of houses with lower grades (1-6), with most grades in this range having negligible representation (close to 0%).The distribution is skewed towards higher grades, emphasizing the prevalence of houses with above-average construction and design quality in the dataset.



## Handling Zero Values in Bedroom & Bathroom 

```{r}
# Function to impute missing values using the median based on non-zero values
impute_nonzero <- function(var) {
  non_zero_values <- as.numeric(var[var != 0])
  if (length(non_zero_values) > 0) {
    imputed_value <- median(non_zero_values)
    var[var == 0] <- imputed_value
  }
  return(var)
}

# Convert the varıables to numeric again
house_data2$bedrooms <- as.numeric(as.character(house_data2$bedrooms))
house_data2$bathrooms <- as.numeric(as.character(house_data2$bathrooms))
house_data2$floors <- as.numeric(as.character(house_data2$floors))

# Apply the imputation function to bedrooms and bathrooms
house_data2$bedrooms <- impute_nonzero(house_data2$bedrooms)
house_data2$bathrooms <- impute_nonzero(house_data2$bathrooms)

```

### Correlation Analysis

```{r}

# Calculate correlation for all variables
cor_matrix <- cor(all_vars)

# Create a heatmap for correlation values
melted_correlation <- melt(cor_matrix)
ggplot(melted_correlation, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1, 1), space = "Lab", 
                       name = "Correlation") +
  geom_text(aes(label = ifelse(abs(value) > 0.5, round(value, 2), "")), vjust = 1,size = 2) + ###we filter the results that are higly correlated to interpret easy  
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_fixed()
```

Initially, we constructed a correlation matrix to discern relationships among variables in the dataset. To enhance interpretability, we applied a filter, selecting correlations with an absolute value greater than 0.5. This focused approach facilitates easier interpretation by highlighting strong correlated variables. 
The choice of the cutoff level for correlation analysis depends on the specific goals of the analysis and the nature of the data. Commonly used cutoff values for correlation coefficients between 0.5-0.7 for moderate correlation and above 0.7 for strong correlation. Eventhough we in the begging choosed 0.7 cutoff , the results showed that The price variable stands out with a strong positive correlation of 0.70 with only the square footage of living space (sqft_living). While this suggests a notable linear relationship between these two factors,we recognized it may be beneficial to also consider variables with moderate correlations to price, as they could provide additional insights into determinants of house prices beyond just living space. So we decreased the cutoff level according that.


According to final results (with 0.5 cutoff):

The Price (dependent variable): Strongly correlated with the square footage of living space (sqft_living)  at 0.70, indicating that larger living spaces tend to command higher prices. And also having moderate correlation with other features such as bathrooms (0.53), sqft_above(0.61), sqft_living15(0.59) and grade(0.67)

The number of bathroom (bathrooms) have strong correlation with only sqft_living(0.75), also having moderate correlation with multiple variables ; price(0.53), bedrooms(0.52), floors (0.5), sqft_above(0.69),sqft_living15 (0.57), grade (0.66),sqft_living15 (0.51)

The square footage of living space (sqft_living) demonstrates strong positive correlations with price (0.70), bathrooms (0.75), sqft_above (0.88), sqft_living15 (0.76), and grade (0.76), highlighting its multifaceted influence on house features and value. and it have moderate correlation with  bedroom (0.58) 

The square footage above ground (sqft_above) has the highest correlation with sqft_living (0.88) and substantial correlations with sqft_living15 (0.73) and grade (0.76), And it have moderate correlation with price (0.61), bathroom (0.69) and floors (0.52) underscoring its significance in determining overall property grades.


The overall grade (grade) exhibits a strong positive correlation with various measures of house size, including sqft_living (0.76), sqft_above (0.76), and sqft_living15 (0.71). Additionally, it shows a moderate correlation with price (0.67) and bathrooms (0.66), suggesting that houses with higher grades tend to be larger, have more bathrooms, and command higher prices.


In conclusion, the correlation analysis has uncovered intricate relationships among various features in the dataset, emphasizing the strong correlation of house prices with the square footage of living space (sqft_living). Additionally, moderate correlations with other features such as bathrooms (0.53), sqft_above (0.61), sqft_living15 (0.59), and grade (0.67) suggest the presence of diverse factors influencing property values, warranting further in-depth analysis in later stages.

Moreover, the identified potential multicollinearity issue highlights the need for careful feature selection to enhance the stability and interpretability of the regression model. Specifically, considering the strong correlations among Sqft_living, Sqft_living15, and Sqft_above, it is advisable to include only one of them in the model to avoid multicollinearity and ensure the model's robustness.

In addition it's crucial to remember that correlation does not imply causation. While these variables are correlated, further analysis and domain knowledge are needed to understand the causal relationships and make informed predictions.

As we move forward, advanced visualizations will serve as valuable tools to unravel these complex relationships, offering a more nuanced understanding of the dynamics shaping the real estate market in King County, Washington State, USA.


### Advanced Visualization Techniques

```{r Scatter plots for continuous variables vs. "price,"}

# Create an empty list to store plots
plots_list <- list()

# Iterate through variables and create scatter plots
for (variable in cont_vars[-1]) {
  # Create scatter plot with regression line
  scatter_plot <- ggplot(house_data2, aes_string(x = variable, y = "price")) +
    geom_point(color = "lightgreen") +
    geom_smooth(method = "lm", se = FALSE, color = "red") +
    geom_encircle(data = house_data2 %>% filter(price > 6000000),
                  color = "red", size = 2, expand = 0.05) +
    labs(title = paste(variable, "vs. Price"), x = variable, y = "Price") +
    theme_minimal()

  # Add the plot to the list
  plots_list[[variable]] <- scatter_plot
}

# Arrange the plots in a grid
advanced_plots <- ggarrange(plotlist = plots_list, ncol = 2, nrow = 2)

# Display the arranged plots
print(advanced_plots)


```

We generate scatter plots for various continuous variables against housing prices, utilizing light green points and red regression lines for visualization. A red circle is incorporated to emphasize observations where housing prices exceed $6,000,000, indicating potential outliers. The scatter plots collectively underscore similarities in the distribution patterns of all continuous variables concerning price. This graphical exploration enhances the understanding of the correlation between each continuous variable and housing prices, with the filter for prices drawing attention to three potential outliers. Identifying and comprehending such outliers is crucial for robust data analysis, aiding in informed decisions regarding their impact on statistical models and subsequent analyses. Further investigation and domain knowledge are typically required to interpret these outliers within the dataset's context.



```{r}
# Create an empty list to store plots
plots_list <- list()

# Iterate through variables and create scatter plots
for (variable in disc_vars) {
  # Create scatter plot with regression line
  scatter_plot <- ggplot(house_data2, aes_string(x = variable, y = "price")) +
    geom_jitter(width = .3, alpha = .3, color = "blue") + # Introduce a noise
    geom_smooth(method = "lm", se = FALSE, color = "red") +
    geom_encircle(data = house_data2 %>% filter(price > 6000000),
                  color = "red", size = 2, expand = 0.05) + # Add an encircling shape for high prices
    geom_encircle(data = house_data2 %>% filter(bedrooms == 33),
                  color = "green", size = 2, expand = 0.05) + # Add an encircling shape for bedrooms == 33
    labs(title = paste(variable, "vs. Price"), x = variable, y = "Price")

  # Add the plot to the list
  plots_list[[variable]] <- scatter_plot
}

# Arrange the plots in a grid
advanced_plots <- ggarrange(plotlist = plots_list, ncol = 2, nrow = 2)

# Display the arranged plots
print(advanced_plots)


```

In this code update, we continued our exploration of the dataset by examining the relationship between housing prices and discrete variables. We introduced noise for a more nuanced view and identified high-priced outliers, visualizing them with encircling shapes.

As an additional step, we focused on the unusual case where the number of bedrooms equals 33. We specifically circled these observations using a distinctive green color. This targeted analysis aims to spotlight and investigate unique patterns and outliers within the data, enhancing our understanding of their impact on housing prices. Our approach reflects an iterative process, adapting visualizations to reveal hidden insights in the dataset.


# Visualize the Distribution of Categorical Variables 


```{r}
# Create an empty list to store plots
plots_list <- list()

# Iterate through categorical variables and create scatter plots
for (variable in cat_vars) {
  # Create scatter plot with regression line, colored points, jitter for density, and circle for high-priced outliers
  scatter_plot <- ggplot(house_data2, aes_string(x = variable, y = "price")) +
    geom_jitter(width = .3, alpha = .3, color= "lightpink") +
    geom_encircle(data = house_data2 %>% filter(price > 6000000),
                  color = "red", size = 2, expand = 0.05) +  # Filter and circle high-priced outliers
    geom_encircle(data = house_data2 %>% filter(bedrooms == 33),
                  color = "green", size = 2, expand = 0.05) +  # Filter and circle 33 bedrooms
    geom_encircle(data = house_data2 %>% filter(grade < 3),
                  color = "blue", size = 2, expand = 0.05) +  # Filter and circle lowest graded once
    labs(title = paste(variable, "vs. Price"), x = variable, y = "Price")

  # Add the plot to the list
  plots_list[[variable]] <- scatter_plot
}

# Arrange the plots in a grid
advanced_plots <- ggarrange(plotlist = plots_list, ncol = 2, nrow = 2)

# Display the arranged plots
print(advanced_plots)


```

In this visualization, we explored various categorical variables in relation to housing prices. Each scatter plot includes a regression line, light pink points with added jitter for better density visualization, and encircling of specific observations. The red circles highlight houses with prices exceeding $6,000,000, signaling potential outliers in the dataset. Additionally, green circles indicate properties with an unusually high number of 33 bedrooms, drawing attention to this unique characteristic. Moreover, blue circles represent homes with a grade lower than 3, suggesting those with the lowest grading. The use of color-coded encircling helps emphasize distinct patterns and potential anomalies in the relationships between categorical variables and housing prices.

The blue circles in the scatter plots indicate houses with the lowest grades (grade < 3). these house is not waterfront , have zero view, and is in a poor condition (condition 1). 

On the contrary, the green circle representing houses with 33 bedrooms stands out with intriguing attributes. Despite boasting an impressive 33 bedrooms, the property lacks waterfront features, has zero views, and exhibits a condition rating of 5, while the grade is higher than 5. While such characteristics are conceivable, the data challenges expectations, especially in terms of the square footage of living space. Conventionally, one might anticipate a residence with an exceptionally high number of bedrooms to accompany a substantial living area. The observed discrepancy between the expected and actual living space raises questions about potential data anomalies or recording errors. An advisable approach involves reassessing or potentially excluding these variables to refine the accuracy and reliability of the analysis. By doing so, the insights derived from the data can better align with realistic expectations and domain knowledge.


```{r}
# Other advanced visualizations for analysis (e.g., hexbin, density plot, etc.)
p<- ggplot(house_data2, aes(x = sqft_living, y = price)) +
  geom_hex(bins = 50) +
  labs(title = 'Hexbin Plot: Housing Price vs. Square Footage of Living Space',
       x = 'Square Footage of Living Space', y = 'Price')  + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Encircle specific data points

# Filter data to select specific data points (sqft_living > 5000 and price > 8000)
house_sel <- house_data2 %>% 
  filter( price > 6000000)

# Add encircling around selected data points with improved aesthetics
p + 
  geom_encircle(data = house_sel, color = "red", size =  1, expand = 0.05, linetype = "dashed") +
    geom_encircle(data = house_data2 %>% filter(bedrooms == 33),
                  color = "green", size = 2, expand = 0.05) +  # Filter and circle 33 bedrooms
    geom_encircle(data = house_data2 %>% filter(grade < 3),
                  color = "blue", size = 2, expand = 0.05) +  # Filter and circle
  theme_bw() +
  scale_x_continuous(labels = scales::comma) +
  scale_y_continuous(labels = scales::dollar_format(scale = 0.001)) +
  theme(axis.text = element_text(size = 10),
        axis.title = element_text(size = 12, face = "bold"),
        plot.title = element_text(size = 14, face = "bold"),
        legend.position = 'top')


```

In this advanced visualization, we employed a hexbin plot to explore the relationship between housing price and square footage of living space. The hexbin plot provides a clear overview of the density of data points, offering insights into the distribution of housing prices concerning living space. We continued the analysis by encircling specific data points of interest. The red dashed circle encompasses houses with prices exceeding $6,000,000. Additionally, we maintained the encircling of the green circle around properties with 33 bedrooms and the blue circle around those with a grade less than 3. These encirclings help to highlight and differentiate distinct subsets within the dataset, providing a nuanced understanding of the data distribution. The refined aesthetics, including color-coded circles and improved line types, enhance the visual appeal and interpretability of the plot. This visualization strategy builds upon the previous stages, offering a comprehensive exploration of the dataset's intricate patterns and outliers.


### Geospatial Visualization
We are planning to use the geographical coordinates that are provided in the dataset.
Create maps to visualize the spatial distribution of house prices in King County and Seattle.
Identify regional patterns and hotspots of high-value properties.


How are house prices distributed geographically in King County?
Are there areas with higher property values?

```{r}

```



## Conclusion
This project aims to provide valuable insights into the factors influencing house sales in King County, Washington, through advanced visualization techniques. The combination of exploratory data analysis will contribute to a comprehensive understanding of the housing market dynamics in the specified region.


##Resources
R Studio and associated libraries (ggplot2, plotly, caret).
Kaggle dataset (provided under CC0: Public Domain).
https://www.kaggle.com/datasets/shivachandel/kc-house-data/data

Access to computational resources for model development.
